---
title: "TermPaper"
output: html_notebook
---
```{r}
library(tidyverse)
library(patchwork)
library(GGally)
```

Source functions and notebooks
```{r,results='hide', include=FALSE,fig.show='hide'}
source("Plant_spectrum_functions.R")
source_rmd("TRY_analysis.Rmd")
source_rmd("GRooT_analysis.Rmd")
source_rmd("GIFT_analysis.Rmd")
#source_rmd("TRY_Gift_GRooT_analysis.Rmd")
```

Load data
```{r}
write_rds(plants_processed, "data/TRY_traits_processed.rds")
```

plots/tables: traits/trait selection, missing traits, TRY pair plot, TRY PCA, GRooT PCA, TGG pair plot, TGG PCoA, TGG woody/non-woody

Missing traits (GRooT)
```{r}
missing_traits(groot) %>% 
  arrange(perc_na) %>% 
  head(15) %>% 
  plot_missing_traits() + labs(title = "GRooT")
```

TRY
  Trait selection
  Merged datasets
  
  Pair plot
```{r}
pair_plots(TRY_pca)
```

  PCA
```{r}
our_pca(TRY3) %>% 
  plot_pcoa(TRY_pca3)
```
  
GrooT
  Trait selection
  Merged datasets
  PCA
GIft ?
  Trait selection
  Merged datasets
  PCoA
TGG
  Trait selection
  Merged datasets
  PCA/PCoA
TGG Woody / non-woody
Patrick’s hypothesis about the Fabaceae
