---
title: "TermPaper"
output: html_notebook
---
```{r}
library(tidyverse)
library(patchwork)
library(GGally)
```

Source functions and notebooks
```{r,results='hide', include=FALSE,fig.show='hide'}
source("Plant_spectrum_functions.R")
source_rmd("TRY_analysis.Rmd")
source_rmd("GRooT_analysis.Rmd")
source_rmd("GIFT_analysis.Rmd")
```


```{r}
theme_set(theme_bw())
```

Load data


plots/tables: traits/trait selection, missing traits, TRY pair plot, TRY PCA, GRooT PCA, TGG pair plot, TGG PCoA, TGG woody/non-woody

Missing traits (GRooT)
```{r}
missing_traits(groot) %>% 
  arrange(perc_na) %>% 
  head(15) %>% 
  plot_missing_traits() + labs(title = "GRooT")
```

TRY
  Trait selection
  Merged datasets
  
  Pair plot
```{r}
pair_plots(TRY_pca)
```

  PCA
```{r}
try_pca <- our_pca(TRY3) #slow!
```


```{r}
plot_pcoa(try_pca, TRY_pca3)
```
  
GrooT
  Trait selection
  Merged datasets
  PCA
  
```{r}
groot_pca <- groot_t4l %>% 
  mutate(species = NA) %>% #the most beatiful line of code ever ;)
  our_pca() #slow!
```


```{r}
plot_pcoa(groot_pca, groot_t4l)
```
GIft ?
  Trait selection
  Merged datasets
  PCoA
TGG
  Trait selection
  Merged datasets
  PCA/PCoA

```{r}
source_rmd("TRY_Gift_GRooT_analysis.Rmd")
```


```{r}
tgg_pcoa_data %>% 
  select(!where(is.character)) %>% 
  #rename_with(~str_wrap(.x, 10)) %>% 
  pair_plots()
```

```{r}
plot_pcoa(tgg_pcoa, try_gift_groot)
```

```{r}
tgg_tree <- try_gift_groot %>% 
  filter(Growth_form_2 == "tree")

tgg_ntree <- try_gift_groot %>% 
  filter(Growth_form_2 != "tree")
```

```{r}
tgg_tree_plot <- tgg_tree %>% 
  pcoa_gower() %>% 
  plot_pcoa(tgg_tree)

tgg_ntree_plot <- tgg_ntree %>% 
  pcoa_gower() %>% 
  plot_pcoa(tgg_ntree)

tgg_tree_plot + tgg_ntree_plot
```


TGG Woody / non-woody
Patrick’s hypothesis about the Fabaceae
