---
title: "Merging TRY, Gift and GRooT"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
---
## Load data
```{r, message=FALSE}
library(tidyverse)
library(StatMatch) # for gower distance
library(vegan) # this is for envfit
library(psych)
library(patchwork)
```

```{r}
try <- readRDS(here::here("data/TRY_traits_processed.rds"))
groot <- readRDS(here::here("data/GRooT_traits_processed.rds"))
gift <- readRDS(here::here("data/GIFT_traits.rds"))
```

# Merge data sets: TRY and gift
```{r}
# see other notebook for this selection
gift_sel_traits <- c("Parasite_1", "Growth_form_2", "Lifecycle_1", "Life_form_1", "Photosynthetic_pathway")
gift_red <- gift %>% 
  select(species, all_of(gift_sel_traits))
```

```{r}
groot_sel_traits <- c("Specific_root_length", "Root_N_concentration", "Mean_Root_diameter")
groot_red <- groot %>% 
  select(species, all_of(groot_sel_traits))
```


```{r}
gift_try <- inner_join(gift_red, try, by = "species")
```


### Compare each column with each other
```{r}
# compares each column with each other to see the pairwise amount of data available
get_amount_data_columnwise <- function(dat, fraction=T, perc=T){
  dat_pres <- dat %>% 
    transmute(across(-species, negate(is.na)))
  ncol <- length(dat_pres)
  amount_data <- map_dfr(seq_along(dat_pres), function(i){
      # each column is compared with every other column only with the diagonal
      row <- rep(NA, ncol) # empty column
      for (j in 1:i){
        present <- sum(dat_pres[i] & dat_pres[j])
        if (fraction){
           present <- (present / nrow(dat_pres) ) %>% 
              round(2)
        }
        row[j] <- present
      }
      
      names(row) <- names(dat_pres)
      if (perc) {row <- row * 100} # convert to percentage
      row
      }
    )
  amount_data %>%  
   mutate(trait = names(dat_pres)) %>% 
    relocate(trait)
}
```


```{r}
(colwise <- gift_try %>% 
  get_amount_data_columnwise)
```
```{r}
trait_order <- names(colwise)
colwise %>%   
  gather("trait2", "data", -trait) %>% 
  mutate(
    trait = factor(trait, rev(trait_order)),
    trait2 = factor(trait2, trait_order)
    ) %>% 
  ggplot(aes(trait, trait2, label=data, fill=data)) +
  geom_label() +
  scale_fill_gradient(low="red", high="green") +
  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
  labs(title= "TRY + GIFT", subtitle = "Percentage of data available for each trait combination")
```

## Remove some traits

by removing SSD we have `898` species

```{r}
gift_try %>% 
  select(-SSD) %>% 
  drop_na
```

by removing also `Life_form_1` we have `1382` species

```{r}
gift_try %>% 
  select(-SSD, -Life_form_1) %>% 
  drop_na
```



# Merge all 3 data sets
```{r}
try_gift_groot_all <- inner_join(groot, try, by= "species") %>% 
  inner_join(gift_red,  by= "species")
```

```{r}
try_gift_groot_all
```

## Compare columns again
```{r}
(colwise2 <- try_gift_groot_all %>% 
  get_amount_data_columnwise)
```

```{r}
trait_order2 <- names(colwise2)
colwise2 %>%   
  gather("trait2", "data", -trait) %>% 
  mutate(
    trait = factor(trait, rev(trait_order2)),
    trait2 = factor(trait2, trait_order2)
    ) %>% 
  ggplot(aes(trait, trait2, label=data, fill=data)) +
  geom_label() +
  scale_fill_gradient(low="red", high="green") +
  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
  labs(title= "All datasets", subtitle = "Percentage of data available for each trait combination")
```

## Remove some traits


```{r}
try_gift_groot_all %>% 
  select(-SSD, -Life_form_1) %>% 
  drop_na()
  
# 240 species
```



```{r}
try_gift_groot_all %>% 
  select(-SSD, -LA) %>% 
  drop_na() 
# 193 species
```
```{r}
try_gift_groot_all %>% 
  select(-SSD, -LA, -Mean_Root_diameter) %>% 
  drop_na() 
# 265 species
```


we decide to remove SSD and LA

```{r}
try_gift_groot <- try_gift_groot_all %>% 
  select(-SSD, -LA) %>% 
  drop_na()
```

```{r}
str(try_gift_groot)
```


```{r}
str(try_gift_groot)
```


```{r}
# tgg stands for Try Gift Groot
tgg_pcoa_data <- try_gift_groot %>% select(-species)
```


```{r}
tgg_dist <- tgg_pcoa_data %>% 
  as.data.frame() %>% 
  gower.dist()
```

```{r}
tgg_pcoa <- cmdscale(tgg_dist, eig=TRUE)
```
```{r}
tgg_pcoa_var <- round(tgg_pcoa$eig/sum(tgg_pcoa$eig)*100, 1)
```


```{r}
tgg_with_pcoa <- try_gift_groot %>% 
  mutate(
    pcoa_ax1=tgg_pcoa$points[,1],
    pcoa_ax2=tgg_pcoa$points[,2]
  )
```


```{r}
ggplot() +
  geom_bar(aes(x=head(tgg_pcoa_var) %>% as.factor(), y = head(tgg_pcoa_var)), stat = "identity")
```


```{r}
tgg_with_pcoa %>% 
ggplot(aes(pcoa_ax1, pcoa_ax2)) +
  geom_point(alpha=.3) +
  xlab(paste("PCoA ax 1 - ", tgg_pcoa_var [1], "%", sep="")) +
  ylab(paste("PCoA ax 2 - ", tgg_pcoa_var [2], "%", sep="")) 
```
Trying to plot the axis and the categories
```{r}
library(vegan)
```


```{r}

tgg_pcoa_data_new_names <- tgg_pcoa_data %>% 
  rename_with(function(col) {paste0(col, "__")}, where(is.character))

tgg_evfit <- envfit(tgg_pcoa, tgg_pcoa_data_new_names)
```
```{r}
tgg_evfit
```
Ask pierre what how we can interpret this r2
```{r}
tgg_evfit$vectors$r
```

```{r}
corr_coef <- 3.5 # This is to make sure that in the plot the arrows looks similar to the points range
tgg_vectors <- 
  tgg_evfit$vectors$arrows %>%
  as.data.frame() %>% 
  add_rownames("trait") %>% 
  mutate(
    Dim1 = Dim1 / corr_coef,
    Dim2 = Dim2 / corr_coef
  )
```


```{r}
add_origins <- function(dat){
  tibble(
    trait = dat$trait,
    Dim1 = 0,
    Dim2 = 0
  ) %>% 
    bind_rows(dat)
}
```


```{r}
tgg_vectors_pos <- tgg_vectors %>% 
  filter(Dim1 > 0) %>% 
  add_origins()
tgg_vectors_neg <- tgg_vectors %>% 
  filter(Dim1 < 0) %>% 
  add_origins()
```


```{r}
ggplot() +
  geom_line(aes(x= Dim1, y = Dim2, colour=trait), data = tgg_vectors_pos, arrow=arrow()) +
  geom_line(aes(x= Dim1, y = Dim2, colour=trait), data = tgg_vectors_neg, arrow=arrow(ends = "first")) +
  geom_label(aes(x=Dim1, y=Dim2, label=trait), data=tgg_vectors)
```

```{r}
ggplot() +
  geom_line(aes(x= Dim1, y = Dim2, colour=trait), data = tgg_vectors_pos, arrow=arrow()) +
  geom_line(aes(x= Dim1, y = Dim2, colour=trait), data = tgg_vectors_neg, arrow=arrow(ends = "first")) +
  geom_label(aes(x=Dim1, y=Dim2, label=trait), data=tgg_vectors) +
  geom_point(aes(pcoa_ax1, pcoa_ax2), alpha=.3, data=tgg_with_pcoa) +
  xlab(paste("PCoA ax 1 - ", tgg_pcoa_var [1], "%", sep="")) +
  ylab(paste("PCoA ax 2 - ", tgg_pcoa_var [2], "%", sep="")) 
```

```{r}
range(tgg_with_pcoa$pcoa_ax1)
```
 #### plotting also categorical traits
 
```{r}
(tgg_factors <- tgg_evfit$factors$centroids %>% 
   as.data.frame() %>% 
   add_rownames("trait_value"))
```

```{r}
tgg_factors <- tgg_factors %>% 
  mutate(
    trait = str_extract(trait_value, ".*(?=__)"), # everything before __
    value = str_extract(trait_value, "(?<=__).*")
  )

```


```{r}
str_extract(tgg_factors$trait, "(?<=__).*")
```

```{r}
tgg_factors %>% 
  select(trait, value) %>% 
  print.data.frame()
```

```{r}
ggplot()+
  geom_point(aes(Dim1, Dim2, col=trait, shape=trait), data = tgg_factors, size=4) +
  geom_text(aes(Dim1, Dim2, label=value), data = tgg_factors, nudge_y = 0.02) +
  xlim(-.22, .22)
```




```{r}
ggplot() +
  geom_line(aes(x= Dim1, y = Dim2, colour=trait), data = tgg_vectors_pos, arrow=arrow()) +
  geom_line(aes(x= Dim1, y = Dim2, colour=trait), data = tgg_vectors_neg, arrow=arrow(ends = "first")) +
  geom_label(aes(x=Dim1, y=Dim2, label=trait), data=tgg_vectors) +
  geom_point(aes(pcoa_ax1, pcoa_ax2), alpha=.3, data=tgg_with_pcoa) +
  
  geom_point(aes(Dim1, Dim2, shape=trait), data = tgg_factors, size=3) +
  geom_text(aes(Dim1, Dim2, label=value), data = tgg_factors, nudge_y = 0.02) +
  
  xlab(paste("PCoA ax 1 - ", tgg_pcoa_var [1], "%", sep="")) +
  ylab(paste("PCoA ax 2 - ", tgg_pcoa_var [2], "%", sep="")) 
```

```{r, warning=F}
library(psych)
pairs.panels(tgg_pcoa_data)
```
#### The H is not not not normal distributed :(

```{r}
#WH doesn't stand for "Wyndham Hotels & Resorts Inc" (first google result) but for Without Height (WH)
tgg_wh <- try_gift_groot_all %>% 
  select(-SSD, -LA, -H) %>% 
  drop_na()
```

```{r}
# tgg stands for Try Gift Groot
tgg_pcoa_data_wh <- tgg_wh %>% select(-species)
```


```{r}
tgg_dist_wh <- tgg_pcoa_data_wh %>% 
  as.data.frame() %>% 
  gower.dist()
```

```{r}
tgg_pcoa_wh <- cmdscale(tgg_dist_wh, eig=TRUE)
```
```{r}
tgg_pcoa_var_wh <- round(tgg_pcoa_wh$eig/sum(tgg_pcoa_wh$eig)*100, 1)
```


```{r}
tgg_with_pcoa_wh <- tgg_wh %>% 
  mutate(
    pcoa_ax1=tgg_pcoa_wh$points[,1],
    pcoa_ax2=tgg_pcoa_wh$points[,2]
  )
```


```{r}
ggplot() +
  geom_bar(aes(x=head(tgg_pcoa_var_wh) %>% as.factor(), y = head(tgg_pcoa_var_wh)), stat = "identity")
```


```{r}
tgg_with_pcoa_wh %>% 
ggplot(aes(pcoa_ax1, pcoa_ax2)) +
  geom_point(alpha=.3) +
  xlab(paste("PCoA ax 1 - ", tgg_pcoa_var_wh [1], "%", sep="")) +
  ylab(paste("PCoA ax 2 - ", tgg_pcoa_var_wh [2], "%", sep="")) 
```
Trying to plot the axis and the categories



```{r}

tgg_pcoa_data_new_names_wh <- tgg_pcoa_data_wh %>% 
  rename_with(function(col) {paste0(col, "__")}, where(is.character))

tgg_evfit_wh <- envfit(tgg_pcoa_wh, tgg_pcoa_data_new_names_wh)
```
```{r}
tgg_evfit_wh
```


```{r}
corr_coef <- 3.5 # This is to make sure that in the plot the arrows looks similar to the points range
tgg_vectors_wh <- 
  tgg_evfit_wh$vectors$arrows %>%
  as.data.frame() %>% 
  add_rownames("trait") %>% 
  mutate(
    Dim1 = Dim1 / corr_coef,
    Dim2 = Dim2 / corr_coef
  )
```

```{r}
tgg_vectors_pos_wh <- tgg_vectors_wh %>% 
  filter(Dim1 > 0) %>% 
  add_origins()
tgg_vectors_neg_wh <- tgg_vectors_wh %>% 
  filter(Dim1 < 0) %>% 
  add_origins()
```


```{r}
(tgg_factors_wh <- tgg_evfit_wh$factors$centroids %>% 
   as.data.frame() %>% 
   add_rownames("trait_value"))
```

```{r}
tgg_factors_wh <- tgg_factors_wh %>% 
  mutate(
    trait = str_extract(trait_value, ".*(?=__)"), # everything before __
    value = str_extract(trait_value, "(?<=__).*")
  )

```


```{r}
ggplot() +
  geom_line(aes(x= Dim1, y = Dim2, colour=trait), data = tgg_vectors_pos_wh, arrow=arrow()) +
  geom_line(aes(x= Dim1, y = Dim2, colour=trait), data = tgg_vectors_neg_wh, arrow=arrow(ends = "first")) +
  geom_label(aes(x=Dim1, y=Dim2, label=trait), data=tgg_vectors_wh) +
  geom_point(aes(pcoa_ax1, pcoa_ax2), alpha=.3, data=tgg_with_pcoa_wh) +
  
  geom_point(aes(Dim1, Dim2, shape=trait), data = tgg_factors_wh, size=3) +
  geom_text(aes(Dim1, Dim2, label=value), data = tgg_factors_wh, nudge_y = 0.02) +
  
  xlab(paste("PCoA ax 1 - ", tgg_pcoa_var_wh [1], "%", sep="")) +
  ylab(paste("PCoA ax 2 - ", tgg_pcoa_var_wh [2], "%", sep="")) 
```
```{r, warning=F}
pairs.panels(tgg_pcoa_data_wh)
```

### comparing distributions

```{r}
compare_dist_trait <- function(trait){
  p1 <- try_gift_groot %>% 
    ggplot(aes_string(x = trait)) +
    geom_histogram(bins=10) +
    labs(title = "Dataset used in PCoA")
  p2 <- try_gift_groot_all %>% 
    ggplot(aes_string(x = trait)) +
    geom_histogram(bins=10) +
    labs(title="Full dataset")
  p1 + p2
}
```


```{r}
compare_dist_trait("H")
```

```{r}
compare_dist_trait("Root_N_concentration")
```

## THE END

```{r}
library(beepr)
```
```{r}
# beep("mario")
# beep("complete")
# beep("fanfare")
# beep("sword") #dasappointing
# beep("wilhelm")
# beep("facebook")
# beep("ready")
# beep("shotgun")
# beep("coin")
# beep("complete")
# beep("treasure")
```

```{r}
celebrations <- function(n=3){
  for (i in 1:n){
    sound <- sample(c("fanfare", "mario", "treasure"), 1)
    beep(sound)
    Sys.sleep(sample(1:5, 1))
  }
}
```

```{r}
#celebrations()
```

